More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-07-25 2016-Weak-FH-Bisim/WeakBisim.tex
First draft introduction

\paragraph{Contribution}
The main interest of our symbolic approach is to define a method to
prove properties directly on open structures, that will be preserved
by any correct instantiation of the holes.

@@

\paragraph{Structure.}

\section{Background and notations}
This section introduces the notations we will use in this article, and  recalls the definition of pNets~\cite{henrio:Forte2016} with an informal semantics  of the pNet constructs.
In general $x_i,y_i$ range over variables.

@@

\subsection{Open Automata: A model of process composition}
\label{sec:OT}

@@

\section{Weak bisimulation}

@@

Below we show the initial transitions of the automaton of our HB
system. The full open-automaton has 17 states, and we have not
computed by hand its full set of transitions. This work would be very
tedious with pen and paper; we are currently developping a software
tool to this end.

>>>>

This article provides a theoretical background that allows us to implement such a verification platform. We build upon the concept of pNets that allowed us to give a behavioural semantics of distributed components and verify the correctness of distributed applications in the past 15 years. pNets  is a
low level semantic framework for expressing the behaviour of various
classes of distributed languages, and as a common internal format for
our tools.  pNets allow the
specification of parameterised hierarchical labelled
transition systems:  labelled transition systems with parameters can be
combined hierarchically.

We develop here a semantics for a model of interacting processes with parameters and holes. 
The main interest of our symbolic approach is to define a method to
prove properties directly on open structures, that will be preserved
by any correct instantiation of the holes.
As a consequence, our model allows us to reason on composition operators as well as on full-size distributed systems. The parametric nature of the model and the properties of compositionability of the equivalence relations thus play a major role here and are the main strengths of  our contribution.

\paragraph{Previous Works and Contribution}
While most of our previous works relied on closed, fully-instantiated semantics~\cite{BBCHM:article2009}, it is only recently that we could design a first version of a  parameterised semantics for pNets with a strong bisimulation equivalence~\cite{henrio:Forte2016}. This article builds open this previous parameterised semantics and provide a cleaner version of the semantics with a slightly simplified formalism and extends it to deal with weak bisimulation. The slight simplification of the formula allowed us to adopt a much cleaner presentation where the theory for open automata, the algebra used to define the semantics of pNets with holes, can be studied independently from its application to pNets. Also, in previous work the study of compositionability was only partial, and in particular the proof that bisimulation is an equivalence is one new contribution of this article and provides a particularly interesting insight on the semantic model we use.
The new formalism allowed us to extend the work and define weak bisimulation  for open automata, which is entirely new. This allows us to define a weak bisimulation equivalence for open pNets with valuable properties of compositionality. 
To summarise, the contribution of this paper are the following
\begin{itemize}
\item The definition of open automata: an algebra of parameterised automata with hole with a strong bisimulation equivalence. This is an adaptation of~\cite{henrio:Forte2016} with an additional property stating that strong bisimulation equivalence is indeed an equivalence relation.
\item A semantics for open pNets expressed as translation to open automata. This is an adaptation of~\cite{henrio:Forte2016} with a complete proof that strong bisimulation is compositional.
\item A theory of weak bisimulation for open automata, and its properties. It relies on the definition of weak open transitions that are derived from transitions of the open automaton by concatenating invisible action transitions with one (visible or not) transition. The precise definition of the concatenation is also a major contribution of this article.
\item A resulting weak bisimulation equivalence for open pNets and a simple static condition on synchronisation vectors inside pNets that is sufficient to ensure that weak bisimulation is compositional.
\item An illustrative example \TODO{TODO}
\end{itemize}

@@

\paragraph{Structure}
This article is organised as follows. Section~\ref{sec:notations} provides the definition of pNets and introduces the notations used in this paper. Section~\ref{sec:OT} defines open automata, i.e. automata with parameters and transitions conditioned by the behaviour of ``holes''; a strong bisimulation equivalence for open automata is also presented in this section. Section~\ref{section:op-semantics} gives the semantics of open pNets expressed as weak automata, and states compositional properties on the strong bisimulation for open pNets. Finally, Section~\ref{sec:weak} defines a weak bisimulation equivalence on open automata and derives weak bisimilarity for pNets, together with properties on compositionality of weak bisimulaiton for open pNets. Finally, Section~\ref{section:conclusion} concludes the paper.

\section{Background and notations}\label{sec:notations}
This section introduces the notations we will use in this article, and  recalls the definition of pNets~\cite{henrio:Forte2016} with an informal semantics  of the pNet constructs.
In general $x_i,y_i$ range over variables.

@@

\section{Open Automata: A model of process composition}
\label{sec:OT}

@@

\section{Weak bisimulation}\label{sec:weak}

@@

Below we show the initial transitions of the automaton of our HB
system. The full open-automaton has 17 states, and we have not
computed by hand its full set of transitions. This work would be very
tedious with pen and paper; we are currently developing a software
tool to this end.
